---
title: "Lab 06 - Sad plots"
author: "Marcus Minko"
date: "02-16-2022"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 

staff <- read_csv("data/instructional-staff.csv")

staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))
```

### Exercise 1

```{r ggplot-line}
staff_long %>%
  ggplot(aes(x = year, y = value, color = faculty_type)) +
  geom_line()

staff_long %>% 
ggplot(aes(x = year, y = value, color = faculty_type, group = faculty_type)) +
      labs(title = "proportion of faculty type", subtitle = "1975-2011", x = "by year", y = "percentage of faculty", color = "type of faculty group") +
     geom_line(size = 2) + theme_minimal()


```
### Exercise 2
```{r reviz-fish}
fisheries <- read_csv("data/fisheries.csv")

fisheries_prop <- fisheries %>% 
 mutate(prop_capture = capture/total, prop_aquaculture = aquaculture/total)

fisheries_prop %>% 
       filter(total > 3000000) %>% 
       ggplot(aes(x = reorder(country, -prop_capture), y = prop_capture)) + geom_bar (stat = 'identity')       + theme_minimal()+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title = "Top fishery producers in the world", x = "Country", y = "Percentage by capture")

```

-I created new variables to translate raw tonnage figures to percentage of total capture and aquaculture. 
-Percentages are standardized and thus better for comparisons when totals vary widely. 
-I filtered for a more manageable subset of countries by tonnage as there are way to many countries in total so it makes sense to select a subset and total makes the most sense as from a production standpoint these countries have the greatest impact or represent the industry in a more substantial way.
-I reorder country by percentage from capture.
-I angled the x-axis labels so they don't overlap.

### Exercise 3
```{r whickham-data}
library(tidyverse) 
library(mosaicData)

data(Whickham)

#observational- no manipulation occuring
#1314 observations, each row is a woman.

ggplot(Whickham, aes(x = outcome)) + geom_bar()
ggplot(Whickham, aes(x = smoker)) + geom_bar()
ggplot(Whickham, aes(x = age)) + geom_histogram (binwidth = 5)

```
...

Add exercise headings as needed.
